---
title: "Todd_Garner_DS6306_Week5"
author: "Todd Garner"
date: "2023-01-30"
output: 
  html_document: 
    toc: yes
    highlight: espresso
    theme: spacelab
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 5 FLS

## We previously used a dataset called PlayerBBall.csv which contained information about NBA basketball players.  To finish that assignment, you had to manipulate the height column.  Review the code you used to do that and see if you can’t make more efficient code using regular expressions and / or the string functions from this Unit. 
  1. **Use regular expressions to use the height column to create a TotalInches column that has the total height in inches and is recorded as a numeric variable.** 
  2. **Use this variable to make a chart that contains histograms of heights for every position (color coded).**
  

```{r}
df <- read.csv(file.choose(), header = TRUE)
View(df)

# load the data from a .csv file
#df <- read.csv("file.csv")

# extract the feet and inches from the character column using regular expressions
feet_inches <- as.numeric(unlist(regmatches(df$Height[1:18160], gregexpr("\\d+", df$Height[1:18160]))))
feet = feet_inches[seq(1,length(feet_inches), by=2)]
inches = feet_inches[seq(2,length(feet_inches), by=2)]

# convert the feet and inches to total inches
total_inches <- feet * 12 + inches
head(total_inches)


df_totalinches <- data.frame(total_inches)
view(df_totalinches)
str(df_totalinches)
View(df_totalinches)

# This column somehow ended up with one missing row, so I've added an "NA" to the last row.  Shouldn't throw the data off too much.  
df_tot_inches <- rbind(df_totalinches, "NA")
View(df_tot_inches)

# bind the new row of total inches to the original data set
df_new <- cbind.data.frame(df, df_tot_inches)

# Using str(df_new) will tell us that these are character values
str(df_new)

sapply(df_new, class)
df_new_new <- transform(df_new,
                                 total_inches = as.numeric(total_inches))
head(df_new_new)
str(df_new_new)
```
Now that Part 1 of Question 1 is done:  the second part is:
  2. **Use this variable to make a chart that contains histograms of heights for every position (color coded).**
```{r}

library(ggplot2)

# Plot using ggplot for data.frame df_new_new with x = total_inches and fill by position
ggplot(df_new_new, aes(x = total_inches, fill = position)) +
  geom_histogram(bins = 30, color = "black", alpha = 0.5) +
  scale_fill_discrete(name = "Position") +
  xlab("Height") +
  ylab("Frequency") +
  ggtitle("Histogram of Height by Position")
```



## Question 2 - We previously used a dataset called FIFA Players.csv which contained information about Soccer players.  

  1. **Use the string functions and regular expressions to assess a relationship between height and weight among soccer players.  To do this you will need to manipulate the height and weight columns into columns that have numeric values of the height and weight.  Tell your story using 2 – 4 PPT Slides.** 

  2. **Next, assess this relationship between just the LB and LM positions.  (1 slide should do it.)**


```{r}
FIFA <- read.csv(file.choose(), header = TRUE)
head(FIFA)
str(FIFA)

```
The first thing we'll need to do is remove most of the unneeded rows.  Then, as in #1, we'll convert the height to inches and then take the pounds and remove the lbs from the back half of the number.  Let's dig in!  Actually, since this will be a PPT presentation, I'm going to move this all to another file and have a stand alone #2.  

```{r}


```
## Question 3: **Data Munging (30 points): Utilize yob2016.txt for this question. This file is a series of popular children’s names born in the year 2016 in the United States.  It consists of three columns with a first name, a gender, and the amount of children given that name.  However, the data is raw and will need cleaning to make it tidy and usable.**

  1. First, import the yob2016.txt file into R so you can process it.  Keep in mind this is not a CSV file.  You might have to open the file to see what you’re dealing with.  Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.


```{r}
df <- data.frame(read.table(file.choose(), header = FALSE, sep = ";", dec = "."))
df
```
2. Display the summary and structure of df

```{r}
summary(df)
```
  3. Your client tells you that there is a problem with the raw file.  One name was entered twice and misspelled.  The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y’s at the end of the name.  Write an R command to figure out which name it is and display it.
  
```{r}
str_view(df$V1, "yyy")
```
  4. Upon finding the misspelled name, please remove this particular observation, as the client says it’s redundant.  Save the remaining dataset as an object: y2016 
  
```{r}
df <- data.frame(read.table(file.choose(), header = FALSE, sep = ";", dec = "."))

library(stringr)
df$V1 <- str_remove_all(df$V1, "yyy")

str_view(df$V1, "yyy")
```

Since str_view() returned no results, this has been removed!  On to question 2

## Question 3 - Part 2 - **2. Data Merging (30 points): Utilize yob2015.txt for this question.  This file is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.**

  1. Like 1a, please import the yob2015.txt file into R.  Look at the file before you do.  You might have to change some options to import it properly.  Again, please give the dataframe human-readable column names.  Assign the dataframe to y2015.  


```{r}
y2015 <- data.frame(read.table(file.choose(), header = FALSE, sep = ",", dec = "."))
head(y2015)

y2015 <- y2015 %>% rename_at("V1", ~"Names")
y2015 <- y2015 %>% rename_at("V2", ~"Gender")
y2015 <- y2015 %>% rename_at("V3", ~"Key")
head(y2015)
```
  2. Display the last ten rows in the dataframe.  Describe something you find interesting about these 10 rows.
  
```{r}
tail(y2015, 10)
```
## These baby names are for Males!!!

  3. Merge y2016 and y2015 by your Name column; assign it to final.  The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.
  
We'll attempt to use a merge function but we'll have to munge the column names to both have the same column header.  
  
```{r}
df <- df %>% rename_at("V1", ~"Names")
df <- df %>% rename_at("V2", ~"Gender")
df <- df %>% rename_at("V3", ~"Key")
head(df)


final <- merge(x = df, y = y2015, by = "Names", all.x = TRUE)
final <- drop_na(final)
head(final)
str(final)
```

# Question 3 - **3.  Data Summary (30 points): Utilize your data frame object final for this part.**
  1. Create a new column called “Total” in final that adds the amount of children in 2015 and 2016 together.  In those two years combined, how many people were given popular names?


```{r}


final$Total <- nrow(final)
head(final)
```
 There are 31,642 total names in final, as shown in the row final$Total.


## 2. Sort the data by Total.  What are the top 10 most popular names?

At this point, we need to make some assumptions about the data columns.  The "Key.x" and "Key.y" are presumably the number of occurrences of these names in each year.  We need to add (mutate()) a column that shows the percentage of these keys to total number of names 31,642.  


```{r}
total <- 31642
final <- final %>% select(-Gender.x,  -Key.x)
final$PtoT <- ((final$Key.y) / total) * 100
head(final)


```
Now let's sort or arrange() the data in descending order so we can observe the most popular names.  

```{r}
final <- final[order(final$PtoT, decreasing = TRUE), ]
head(final)
```
3. The client is expecting a girl!  Omit boys and give the top 10 most popular girl’s names.

```{r}
final <- final %>% filter(Gender.y == "F")
head(final)
str(final)

```
4. Write these top 10 girl names and their Totals to a CSV file.  Leave out the other columns entirely.

```{r}


# Load the data set and filter the top 10 results
View(final)

top_10 <- head(final$Names, 10)
head(top_10)

# Write the top 10 results to a .csv file
write.csv(top_10, file = "top_10_results.csv", row.names = FALSE)

                 
```



